Time stability in human interaction networks

### 2015-06-17 paper.tex
realocado apendice para metodologia

\begin{abstract}
In this study, we demonstrate a remarkably stable activity in human interaction networks. The activity along time and topology evolution were investigated in four e-mail lists by considering window sizes from 50 to 10,000 messages, which were made to slide and generate snapshots of the network in a timeline. Furthermore, the activity in terms of seconds, minutes, hours, days and months, is practically the same for all lists. The activity of participants followed the expected scale-free behavior, thus allowing us to establish three classes of vertices by comparing with the Erd\"os-R\'enyi model, namely hubs, intermediary and peripheral vertices. The relative size of these three sectors did not vary with time and was essentially the same for all e-mail lists. Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and the remainder are peripheral vertices. The metrics that contribute most to the dispersion of participants in the topological measures space were centrality measurements (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient. Similar results for the distribution of participants in the three categories and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participa.br. Consistent with expectations driven from the literature, these properties may be general for human interaction networks, which has important implications in establishing a typology based on objective, quantitative criteria.
\end{abstract}

@@

\section{Introduction}\label{sec:into}
Studies on human interaction networks have started long before modern computers, dating back to the nineteenth century, while the foundation of
social network analysis is generally attributed to the psychiatrist Jacob Moreno in mid twentieth century~\cite{newmanBook}. With the increasing availability of data related to human interactions, research on these networks has grown continuously. Contributions can now be found in a variety of fields, from social sciences and humanities~\cite{latour2013} to computer science~\cite{bird} and physics~\cite{barabasiHumanDyn,newmanFriendship}, given the multidisciplinary nature of the topic. One of the approaches from an exact science perspective is to represent interaction networks as complex networks~\cite{barabasiHumanDyn,newmanFriendship}, with which 
several features of human interaction have been revealed. For example, the topology of human interaction networks exhibits a scale-free trace, which points to the existence of a small number of highly connected hubs and a large number of poorly connected nodes. The dynamics of complex networks representing human interaction has also been addressed ~\cite{barabasiEvo,newmanEvolving}, but only to a limited extent, since research is normally focused on a particular metric or task, such as accessibility or community detection~\cite{access,newmanModularity}.

@@

The paper is organized as follows. Section~\ref{sec:related} describes related work, while details of the data and methods of analysis are given in  Section~\ref{sec:data} and Section~\ref{sec:carac}. Section~\ref{sec:results} brings the results and discussion, leading to Section~\ref{sec:conc} for conclusions. Details of how to access the data are given in the Appendix, while subsidiary results from the email lists and of networks from Twitter, Facebook and Participa.br are given in the Supporting Information.

\subsection{Related work}\label{sec:related}
Research on network evolution often considers solely network growth, in which there is a monotonic increase in the number of events considered~\cite{barabasiEvo}. Exceptions are reported in this section, with emphasis on those more closely related to the present article.

@@

\begin{table}
	\centering
	\caption{Columns $date_1$ and $date_M$ have dates of first and last messages from the 20,000 messages considered in each email list.
		$N$ is the number of participants (number of different email addresses).
		$\Gamma$ is the number of threads (count of messages without antecedent).
		$\overline{M}$ is the number of messages missing in the 20,000 collection, $100\frac{23}{20000}=0.115$ percent in the worst case.
	A relation holds for all these lists: within a same number of messages, as the number of participants increases, the number of threads decreases.}
	\label{tab:genLists}
	\begin{tabular}{|l|c|c|c|c|c|}\hline
		list & $date_1$ & $date_{M}$    & $N$  & $\Gamma$ & $\overline{M}$ \\\hline
		\input{tables/tab1Geral}
	\end{tabular}
\end{table}

@@

The simplification of all the compound possibilities to the small set listed above can be formalized in strict mathematical terms, but this was considered out of the scope for current interests. It is worth noting that the exclusivist cascade is the same sectioning of an inclusivist cascade from periphery to hubs, but with inverted order of sectors precedence. These compound criteria can be used to examine network sections considering all degrees and strengths and are specially useful in the case of a low number of messages, such as in Section~\ref{si:frac} of the Supporting Information.

@@

@@

In order to verify the possible generality of these findings, we obtained the Erd\"os sectors of 12 networks from Facebook, Twitter and Participa.br. The results are given in Tables~\ref{tab:secE} in the Supporting Information, which indicate that the percentages of hubs, intermediary and peripheral nodes are essentially the same as for the email lists.

@@

A systematic study of the activity of participants belonging to the three distinct Erd\"os sectors indicated simple patterns for hubs and peripheral vertices, while the network structure was governed by the intermediary vertices. These properties were shared by all email lists and were time-independent, being consistent with the literature. Moreover, both the distribution of Erd\"os sectors and the contribution from the metrics to the PCA were found to apply to networks from Facebook, Twitter and Participa.br. We may therefore consider the classification of agents into Erd\"os sectors as leading to a human typology which bridges between exact sciences, with objective procedures for the classification, and human sciences, where there is a legacy in the observation of human types.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section*{APPENDIX: Data and scripts}\label{scripts}
Messages were downloaded from the GMANE public database~\cite{GMANEwikipedia}. Data annotated from Facebook and Twitter are in a public repository~\cite{fbtwData}. Data from Participa.br was used from the linked data/semantic web RDF triples reported in~\cite{opa} and available in~\cite{datahub}.
All routines necessary to achieve the results reported in this article, including all tables and figure of the Supporting Information, are available through a public domain Python package and an open Git repository~\cite{gmanePack}. Data from social networks used in this study were gathered and used within the Anthropological Physics principles~\cite{anPhy}.

%\nocite{*}
\bibliography{paper}% Produces the bibliography via BibTeX.

>>>>

\begin{abstract}
In this study, we demonstrate a remarkably stable activity in human interaction networks. The activity along time and topology evolution were investigated in four e-mail lists by considering window sizes from 50 to 10,000 messages, which were made to slide and generate snapshots of the network in a timeline. Furthermore, the activity in terms of seconds, minutes, hours, days and months, is practically the same for all lists. The activity of participants followed the expected scale-free behavior, thus allowing us to establish three classes of vertices by comparing with the Erd\"os-R\'enyi model, namely hubs, intermediary and peripheral vertices. The relative size of these three sectors did not vary with time and was essentially the same for all e-mail lists. Typically, 3-12\% of the vertices are hubs, 15-45\% are intermediary and the remainder are peripheral vertices. The metrics that contribute most to the dispersion of participants in the topological measures space were centrality measurements (degree, strength and betweenness), followed by symmetry-related metrics and then clustering coefficient. Similar results for the distribution of participants in the three categories and for the relative importance of the topological metrics were obtained for 12 additional networks from Facebook, Twitter and Participa.br. Consistent with expectations from the literature, these properties may be general for human interaction networks, which has important implications in establishing a typology based on objective, quantitative criteria.
\end{abstract}

@@

\section{Introduction}\label{sec:into}
Studies on human interaction networks have started long before modern computers, dating back to the nineteenth century, while the foundation of
social network analysis is generally attributed to the psychiatrist Jacob Moreno in mid twentieth century~\cite{newmanBook}. With the increasing availability of data related to human interactions, research on these networks has grown continuously. Contributions can now be found in a variety of fields, from social sciences and humanities~\cite{latour2013} to computer science~\cite{bird} and physics~\cite{barabasiHumanDyn,newmanFriendship}, given the multidisciplinary nature of the topic. One of the approaches from an exact science perspective is to represent interaction networks as complex networks~\cite{barabasiHumanDyn,newmanFriendship}, with which 
several features of human interaction have been revealed. For example, the topology of human interaction networks exhibits a scale-free trace, which points to the existence of a small number of highly connected hubs and a large number of poorly connected nodes. The dynamics of complex networks representing human interaction has also been addressed~\cite{barabasiEvo,newmanEvolving}, but only to a limited extent, since research is normally focused on a particular metric or task, such as accessibility or community detection~\cite{access,newmanModularity}.

@@

The paper is organized as follows. Section~\ref{sec:related} describes related work, while details of the data and methods of analysis are given in  Section~\ref{sec:data} and Section~\ref{sec:carac}. Section~\ref{sec:results} brings the results and discussion, leading to Section~\ref{sec:conc} for conclusions.
Subsidiary results from the email lists and of networks from Twitter, Facebook and Participa.br are given in the Supporting Information.

\subsection{Related work}\label{sec:related}
Research on network evolution often considers solely network growth, in which there is a monotonic increase in the number of events~\cite{barabasiEvo}. Exceptions are reported in this section, with emphasis on those more closely related to the present article.

@@

\begin{table}
	\centering
	\caption{Columns $date_1$ and $date_M$ have dates of first and last messages from the 20,000 messages considered in each email list.
		$N$ is the number of participants (number of different email addresses).
		$\Gamma$ is the number of threads (count of messages without antecedent).
		$\overline{M}$ is the number of messages missing in the 20,000 collection, $100\frac{23}{20000}=0.115$ percent in the worst case.
	}
	\label{tab:genLists}
	\begin{tabular}{|l|c|c|c|c|c|}\hline
		list & $date_1$ & $date_{M}$    & $N$  & $\Gamma$ & $\overline{M}$ \\\hline
		\input{tables/tab1Geral}
	\end{tabular}
\end{table}

@@

The simplification of all the compound possibilities to the small set listed above can be formalized in strict mathematical terms, but this was considered out of the scope for current interests. It is worth noting that the exclusivist cascade is the same sectioning of an inclusivist cascade from periphery to hubs, but with inverted order of sectors precedence. These compound criteria can be used to examine network sections considering all degrees and strengths and are especially useful in the case of a low number of messages, such as in Section~\ref{si:frac} of the Supporting Information.

@@

\subsection{Data and scripts availability}\label{scripts}
As a way to share routines with the scientific community and the whole of the society,
all the software necessary to achieve the results reported in this article,
including all tables and figure of the Supporting Information,
are available through a public domain Python package and an open Git repository~\cite{gmanePack}.

Data from social networks used in this study were gathered and used within the Anthropological Physics principles~\cite{anPhy}.
All data used are also publicly accessible, either because they were already in public domain or because
we published our own annotations.
Messages were downloaded from the GMANE public database~\cite{GMANEwikipedia}.
Data annotated from Facebook and Twitter are in a public repository~\cite{fbtwData}.
Data from Participa.br was used from the linked data/semantic web RDF triples
reported in~\cite{opa} and available in~\cite{datahub}.

This is a way to enhance the reliability of the methods,
of algorithmic routines, of data consistency,
and of the results themselves.
Also, we believe that this practice raises the scientific contribution of the work,
handing not only the framework and the results, 
but the exact data and processes that render them~\cite{openSci}.

@@

In order to verify the possible generality of these findings, we obtained the Erd\"os sectors of 12 networks from Facebook, Twitter and Participa.br. The results are given in Table~\ref{tab:secE} in the Supporting Information, which indicate that the percentages of hubs, intermediary and peripheral nodes are essentially the same as for the email lists.

@@

A systematic study of the activity of participants belonging to the three distinct Erd\"os sectors indicated simple patterns for hubs and peripheral vertices, while the network structure was governed by the intermediary vertices. These properties were shared by all email lists and were time-independent, being consistent with the literature. Moreover, both the distribution of Erd\"os sectors and the contribution from the metrics to the PCA were found to apply to networks from Facebook, Twitter and Participa.br. We may therefore consider the classification of agents into Erd\"os sectors as leading to a human typology which bridges exact sciences, with objective procedures for the classification, and human sciences, where there is a legacy in the observation of human types.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\nocite{*}
\bibliography{paper}% Produces the bibliography via BibTeX.
